ME780 Assignment 2

### 2016-04-03 assignments/assignment_2/report.tex
last round of edits

\section{Motion Model of a Bicycle}
Define the bicycle motion model for an Ackermann steered robot as shown in the configuration below(and presented in the motion modelling lecture notes). 
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{images/bicycle_model.png}
	\caption{Bicycle in World Frame}
	\label{fig:bicycle}
\end{figure}

@@

Develop a carrot following controller.  The way this controller works is you define a carrot, or point on the desired trajectory line a fixed distance, $r$, ahead of the closest point on the line to the current robot position.

@@

A carrot controller is a controller that has a carrot point with a fixed distance $r$ on the trajectory line, ahead of the closest point on the trajectory line to the current robot position.

\subsection{Closest Point on the Trajectory Line to the Robot}
The closest point between the robot's position and the trajectory line can be calculated by first calculating the projection of the vector formed by the current location and the desired

@@

\newpage
\subsection{Calculation of the Carrot Point}
Once the closest point on the trajectory line to the robot is found, calculating a carrot point along the trajectory line is the closest point on the trajectory line $p_{\text{closest}}$ plus the product of look ahead distance $r$ times the unit vector of the trajectory line $b$.

\begin{equation}
	\text{carrot} = p_{\text{closest}} + r \dfrac{\vec{b}}{|\vec{b}|}
\end{equation}

\subsection{Carrot Controller}
We implemented a proportional (P) carrot controller and traced a rectangle of $20m$ by $5m$ at $3ms^{-1}$. This is the resulting motion of the bicycle (See Fig~\ref{fig:bicycle_carrot}), we used a proportional gain $k_{p} = 2$ and look ahead distance $r$ at $2$, $0.2$ and $5$.

From Fig~\ref{fig:bicycle_carrot} we can see that different look ahead distance $r$ impacts the way the robot traces a rectangle. If the look ahead distance is too low (i.e. Fig~\ref{fig:bicycle_carrot_bad} where $r = 0.2$), the carrot will be a very small distance ahead of the robot, making small adjustments unlikely thereby the angle of the steering oscillates between $-30$ degrees to $30$ degrees (the max steering angle).

If the carrot is at a medium look ahead distance away from the robot  (i.e. Fig~\ref{fig:bicycle_carrot_ok} where $r = 2$), the robot tends to be quite agreessive in compensating for overshooting the trajectory. As we increase the look ahead distance $r$ however, the correction angle is dampened, causing a less agreessive correction, but longer time to return to the trajectory line.

@@

\newpage
\section{IGVC Planner}
\label{sec:igvc_planner}
We were given a map of the Intelligent Ground Vehicle Competition in the form of a $926 \times 716$ element grid, where each cell
represents a 10 cm X 10 cm area of the course, resulting in a 92.6 m X 71.6 m environment.

@@

\newpage
\subsection{Wavefront Planner}
\label{subsec:wavefront}
Since the map given represents an occupancy grid of free space and obstacles, our initial approach to the planning portion was to apply the wavefront a path that finds an optional path between start to goal point. Fig~\ref{fig:wavefront} highlights the results of the wavefront planner and highlights the resulting path found by gradient decent. 
\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{images/wavefront_igvc.eps}
	\caption{Wavefront IGVC Map}
	\label{fig:wavefront}
\end{figure}

@@

Lastly, for this large map, the run time was over 40 seconds without any collision checking procedures and therefore this approach was abandoned in favour of a Probabilistic Roadmap Planner (PRM).

@@

\section{Limitations and Future Improvements to the PRM Planner}
\subsection{Limitations}

>>>>

\section{Motion Model of a Bicycle}
\textit{Define the bicycle motion model for an Ackermann steered robot as shown in the configuration below(and presented in the motion modelling lecture notes). }
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{images/bicycle_model.png}
	\caption{Bicycle in World Frame}
	\label{fig:bicycle}
\end{figure}

@@

\textit{Develop a carrot following controller.  The way this controller works is you define a carrot, or point on the desired trajectory line a fixed distance, $r$, ahead of the closest point on the line to the current robot position.}

@@

\subsection{Closest Point on the Trajectory Line to the Robot}
The first step in implementing the the carrot controller for the outlined problem is to first determine a way to calculate the correct location of the carrot point given the robot's current position. To do we can use the projection equation between two vectors.

@@

Next we need to determine where to place the carrot point. Having already found the closest point on the trajectory line $p_{\text{closest}}$, the carrot is simply calculated by adding the look ahead distance $r$ times the unit vector of the trajectory line $b$ to the value of $p_{\text{closest}}$.
\begin{equation}
	\text{carrot} = p_{\text{closest}} + r \dfrac{\vec{b}}{|\vec{b}|}
\end{equation}

\subsection{Carrot Controller Implementation}
We implemented a proportional (P) carrot controller and traced a rectangle of $20m$ by $5m$ at $3ms^{-1}$. This is the resulting motion of the bicycle (See Fig~\ref{fig:bicycle_carrot}), we used a proportional gain $k_{p} = 2$ and tested look ahead distances $r$ at $2$, $0.2$ and $5$.

From Fig~\ref{fig:bicycle_carrot} we can see that different look ahead distance $r$ greatly impact the way the robot traces a rectangle. If the look ahead distance is too low (i.e. Fig~\ref{fig:bicycle_carrot_bad} where $r = 0.2$), the carrot is always very small distance ahead to the robot which in turn causes the angle between the carrot's location and the robot to be relatively large. This combined with the course time step of $dt = 0.1s$ cause the robot to overshoot and cross over the desired path line at each time step, causing a large error in the heading which then causes it to overshoot again, leading to the oscillatory motion observed in Fig~\ref{fig:bicycle_carrot_bad}. This problem may be avoided if one was to use a PD controller, reduce the P gains or decrease the time step.

When the carrot was set to a look ahead distance of $r = 5m$ we found that robot had trouble hitting a way point and we had to increase the way point reached threshold from $0.2 m $ to $0.5m$ to compensate for this problem. The way point reached threshold represented the circle that the robot must enter in order for the carrot controller to consider the way point to have been reached. This difficulty of hitting way points, along with larger overshoots compared to the carrot distances were $r= 0.2$ and $r = 2m$ are caused by the large distance between the carrot and the robot which reduces the calculated angle between the robot and the carrot at each time step, making it less sensitive to heading errors and in turn dampening the turning rate of the robot.

Overall we found that a lookahead distance of $r = 2m$ provided the best results for the time steps and proportional gains provided. We were able to use a detection radius of 0.2 at each way point to detect if the robot had reached the way point goal. Moreover, after passing a way point and slightly overshooting, the robot's trajectory matched the on the next path segment within approximately $5m$ of last way point compared to the runs where $r = 5m$ in which case the robot took over $10m$ to pass directly on the path segments.

@@

\newpage
\section{IGVC Planner}
\label{sec:igvc_planner}
\textit{Using the map provided with the homework (IGVCmap.mand IGVCmap.jpg), identify a planning strategy that can be used in conjunction with the carrot planner of part 2)to navigate the Intelligent Ground Vehicle Competition auto-navigation course.}

The provided map of the Intelligent Ground Vehicle Competition in the form of a $926 \times 716$ element grid, where each cell
represents a 10 cm X 10 cm area of the course, resulting in a 92.6 m X 71.6 m environment.

@@

\subsection{Wavefront Planner}
\label{subsec:wavefront}
Since the map given represents an occupancy grid of free space and obstacles, our initial approach to the planning portion was to apply the wavefront a path that finds an optional path between start to goal point. Fig~\ref{fig:wavefront} highlights the results of the wavefront planner and highlights the resulting path found by gradient decent. 
\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\textwidth]{images/wavefront_igvc.eps}
	\caption{Wavefront IGVC Map}
	\label{fig:wavefront}
\end{figure}

@@

Lastly, for this large map, the run time was over 1 minute without any collision checking procedures and therefore this approach was abandoned in favour of a Probabilistic Roadmap Planner (PRM).

@@

\section{Limitations and Future Improvements to the PRM Planner}
